---
title: "MP Genome Comparison"
author: "Maxwell Katsikas, Joel Pantusso"
date: '2022-04-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
```

# Reading in strain annotations #
```{r}
strain1 <- read.csv("SRR2135844.csv") %>% 
  mutate(strain = 1) %>% 
  filter(Feature != "gene") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain) %>% 
  mutate(Product = '',
         ID = '',
         RefSeq = '')

strain1_compare <- read.csv("SRR2135844.csv") %>% 
  mutate(strain = 1) %>% 
  filter(Feature != "CDS") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain)

strain2 <- read.csv("SRR2135847.csv")%>% 
  mutate(strain = 2) %>% 
  filter(Feature != "gene") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain) %>% 
  mutate(Product = '',
         ID = '',
         RefSeq = '')

strain2_compare <- read.csv("SRR2135847.csv") %>% 
  mutate(strain = 2) %>% 
  filter(Feature != "CDS") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain)

strain3 <- read.csv("SRR2135848.csv")%>% 
  mutate(strain = 3) %>% 
  filter(Feature != "gene") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain) %>% 
  mutate(Product = '',
         ID = '',
         RefSeq = '')

strain3_compare <- read.csv("SRR2135848.csv") %>% 
  mutate(strain = 3) %>% 
  filter(Feature != "CDS") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain)

strain4 <- read.csv("SRR2135850.csv")%>% 
  mutate(strain = 4) %>% 
  filter(Feature != "gene") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain) %>% 
  mutate(Product = '',
         ID = '',
         RefSeq = '')

strain4_compare <- read.csv("SRR2135850.csv") %>% 
  mutate(strain = 4) %>% 
  filter(Feature != "CDS") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain)

strain5 <- read.csv("SRR2135852.csv")%>% 
  mutate(strain = 5) %>% 
  filter(Feature != "gene") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain) %>% 
  mutate(Product = '',
         ID = '',
         RefSeq = '')

strain5_compare <- read.csv("SRR2135852.csv") %>% 
  mutate(strain = 5) %>% 
  filter(Feature != "CDS") %>% 
  select(Sequence.H1A1.H1, Feature, Start, End, Strand, Attributes, strain)
```

# Creating ID attribute function #
```{r}
Attribute_ID <- function(x) {
  Attributes_list <- str_split(x$Attributes, ";")
  Attributes <- do.call(rbind, Attributes_list)
  ID <- str_detect(Attributes, "ID=")
  Attributes_DF <- data.frame(Attributes) %>%
    rename("ID" = X1) %>%
    select(ID)
  x <- x %>%
    mutate(ID = cbind(Attributes_DF))
}
```

# Creating refseq attribute function #
```{r}
get_refseq <- function(x) {
  Attributes_list <- str_split(x$Attributes, ";")
  Attributes <- do.call(rbind, Attributes_list)
  for (row in 1:nrow(Attributes)) {
    start_loc <- str_locate(Attributes[row], "RefSeq=")
    end_loc <- start_loc + 21
    refseq <- x$Attributes[row][start_loc:end_loc]
  x$RefSeq[row] <- refseq
  }
}
```

# Creating gene comparison function #
```{r}
gene_comparison <- function(strain) {
  num_gene <- 0
  num_trna <- 0

  
  for (row in 1:nrow(strain)) {
    
    if (strain$Feature[row] == "gene") {
      strain$gene[row] <-  1
      num_gene <- num_gene +1
    } else {
      strain$gene[row] <-  0
    }
    
    if (strain$Feature[row] == "tRNA") {
      strain$trna[row] <-  1
      num_trna <- num_trna + 1
    } else {
      strain$trna[row] <-  0
    }
  }
    
  percent_gene <- num_gene / nrow(strain)
  percent_trna <- num_trna / nrow(strain)
  strain <- strain %>% 
    mutate(num_gene <- num_gene) %>% 
    mutate(num_trna <- num_trna) %>% 
    mutate(percent_gene <- percent_gene) %>% 
    mutate(percent_trna <- percent_trna)
}
```


# Applying gene comparison function #
```{r}
strain1_compare <- gene_comparison(strain1_compare)
strain2_compare <- gene_comparison(strain2_compare)
strain3_compare <- gene_comparison(strain3_compare)
strain4_compare <- gene_comparison(strain4_compare)
strain5_compare <- gene_comparison(strain5_compare)
```


# Merging strain data frames #
```{r}
strain_compare <- rbind(strain1_compare, strain2_compare)
strain_compare <- rbind(strain_compare, strain3_compare)
strain_compare <- rbind(strain_compare, strain4_compare)
strain_compare <- rbind(strain_compare, strain5_compare) %>% 
  group_by(strain) %>% 
  summarize(num_gene = sum(gene),
            num_trna = sum(trna))
```

# Adding ID Attribute Column #
```{r, warning=FALSE}
strain1_ID <- Attribute_ID(strain1)
strain2_ID <- Attribute_ID(strain2)
strain3_ID <- Attribute_ID(strain3)
strain4_ID <- Attribute_ID(strain4)
strain5_ID <- Attribute_ID(strain5)
```


# Comparing what ID's are shared across strains #
```{r}
shared_IDs <- intersect(strain1_ID$ID, strain2_ID$ID)
shared_IDs <- intersect(shared_IDs, strain3_ID$ID)
shared_IDs <- intersect(shared_IDs, strain4_ID$ID)
shared_IDs <- intersect(shared_IDs, strain5_ID$ID)
```


# Adding RefSeq Attribute Column #
```{r}
#strain1_RefSeq <- get_refseq(strain1_ID)
```


# Plot of Number of Genes Annotated #
```{r}
genes_plot <- ggplot(data=strain_compare) +
  geom_col(aes(x=strain, y=num_gene), fill="darkgrey") +
  labs(title="Genes Annotated Across Five Strains of Mycoplasma pneumoniae", 
       caption="Source: NCBI") +
  coord_cartesian(ylim = c(650, 750)) +
  annotate("text", label="Strain M547", x= 1, y=700, color="black") +
  xlab("Strain") +
  ylab("Number of Genes Annotated") +
  theme_economist() +
  theme(plot.title = element_text(size=14))
```

```{r}
ggsave(filename="gene_plot.png", plot = genes_plot, path = "C:/Users/maxka/OneDrive/Documents/BIO485")
genes_plot
```

# Plot of Number of tRNA Annotated #
```{r}
tRNA_plot <- ggplot(data=strain_compare) +
  geom_col(aes(x=strain, y=num_trna), fill="darkgrey") +
  labs(title="tRNA Segments Annotated Across Five Strains of Mycoplasma pneumoniae", 
       caption="Source: NCBI") +
  coord_cartesian(ylim = c(30, 38)) +
  scale_y_continuous(breaks=c(30, 32, 34, 36, 38)) +
  annotate("text", label="Strain M547", x= 1, y=35, color="black") +
  xlab("Strain") +
  ylab("Number of tRNA Annotated") +
  theme_economist() +
  theme(plot.title = element_text(size=12.5))
```

```{r}
ggsave(filename="tRNA_plot.png", plot = tRNA_plot, path = "C:/Users/maxka/OneDrive/Documents/BIO485")
tRNA_plot
```

```{r, warning=FALSE}
M547 <- read.table(text=as.character(strain1$Attributes), sep=";", fill=TRUE,
                   as.is=TRUE) %>% 
  select(V1, V4, V7, V8) %>% 
  rename("ID" = V1, "Ontology" = V4, "Product" = V7, "Protein_ID" = V8)

strain2_att <- read.table(text=as.character(strain2$Attributes), sep=";", fill=TRUE,
                   as.is=TRUE) %>% 
  select(V1, V4, V7, V8) %>% 
  rename("ID" = V1, "Ontology" = V4, "Product" = V7, "Protein_ID" = V8)

strain3_att <- read.table(text=as.character(strain3$Attributes), sep=";", fill=TRUE,
                   as.is=TRUE) %>% 
  select(V1, V4, V7, V8) %>% 
  rename("ID" = V1, "Ontology" = V4, "Product" = V7, "Protein_ID" = V8)

strain4_att <- read.table(text=as.character(strain4$Attributes), sep=";", fill=TRUE,
                   as.is=TRUE) %>% 
  select(V1, V4, V7, V8) %>% 
  rename("ID" = V1, "Ontology" = V4, "Product" = V7, "Protein_ID" = V8)

strain5_att <- read.table(text=as.character(strain5$Attributes), sep=";", fill=TRUE,
                   as.is=TRUE) %>% 
  select(V1, V4, V7, V8) %>% 
  rename("ID" = V1, "Ontology" = V4, "Product" = V7, "Protein_ID" = V8)
```

```{r}
shared_products <- intersect(M547$Product, strain2_att$Product)
shared_products <- intersect(shared_products, strain3_att$Product)
shared_products <- intersect(shared_products, strain4_att$Product)
shared_products <- intersect(shared_products, strain5_att$Product)
shared_products
```

```{r}
shared_ontology <- intersect(M547$Ontology, strain2_att$Ontology)
shared_ontology <- intersect(shared_ontology, strain3_att$Ontology)
shared_ontology <- intersect(shared_ontology, strain4_att$Ontology)
shared_ontology <- intersect(shared_ontology, strain5_att$Ontology)
shared_ontology
```
